---
title: "data_manipulation"
Author: "Xue Zhang"
Date: "2023-09-19"
output: github_document
---



```{r}
library(tidyverse)
```


Let's import the FAS_litters.csv` csv using a relative path.



```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df = 
  janitor::clean_names(pups_df)
```


```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, gd0_weight, group)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, group, starts_with("pups"))

select(litters_df, -litter_number)

select(litters_df, -starts_with("gd"))

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())

select(litters_df, gd0_weight, everything())
```

```{r}
rename(litters_df, litter_id = litter_number)

relocate(litters_df, litter_number)
```

`rename` is a short hand 


`select` s `pull`

```{r}
select(litters_df, group)

pull(litters_df, group)
```

If need something to get off from the data frame , use `pull`.


pracftice with pups.csv.

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

get rid of rows using `filter()`

filter something exist inside the data frame

== to check this thing if turns out to be true

```{r}
filter(litters_df, group == "Con7")

filter(litters_df, group == "Mod8")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, gd0_weight <= 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20 )
```

what about missing rows

drop missing values, "NA"

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```


## `mutate`

this is used to add or change variables

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```






